Measurement and Probability in Fuzzy Quantum Theories

### 2017-01-22 prop.tex
prop.tex \section{Increasing the Measurement Resource} Updating...

Although quantum IVPMs in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-4-value} have empty cores, we might
still think that the probability measure~$\bar{\mu}$ is induced
by some states. In example~\ref{ex:three-dimensional-three-value},
the probability measure~$\bar{\mu}$ might be induced by the state
$\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$.
In example~\ref{ex:three-dimensional-4-value}, the probability measure~$\bar{\mu}$
might be induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$.
In general, each two real 1-dimensional projectors~$\proj{\psi_{0}}$
and $\proj{\psi_{1}}$ mapping to impossible suggests $\bar{\mu}$
might be induced by the pure state~$\ket{\psi_{0}}\times\ket{\psi_{1}}$
in the Hilbert space of dimension 3.\footnote{Notice that we only discuss only discuss real 1-dimensional projectors
in the Hilbert space of dimension 3. In general, we could discuss
all 1-dimensional projectors in any dimension. However, restricting
to $\mathbb{R}^{3}$ enable us to use cross product which will simplify
our discussion a lot, and $\mathbb{R}^{3}$ is easy to visualize and
imagine. Also, if a quantum IVPM~$\bar{\mu}$ has an empty core when
restricting to $\mathbb{R}^{3}$, $\bar{\mu}$ has an empty core on
the original space as well. So restricting on real 1-dimensional projectors
still provides valid results.

About the cross product, even if $\ket{\psi_{0}}$ and $\ket{\psi_{1}}$
are normalized, when they are not orthogonal, $\ket{\psi_{0}}\times\ket{\psi_{1}}$
is not normalized. We will use $\ket{\psi_{0}}\times\ket{\psi_{1}}$
and the normalized $\ket{\psi_{0}}\times\ket{\psi_{1}}$ interchanged
when there is no confusion.}

@@

\begin{lemma}A interval-valued frame function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the following properties:
\begin{itemize}
\item $\bar{f}\left(\lambda\ket{\psi}\right)=\bar{f}\left(\ket{\psi}\right)$
for any $\ket{\psi}\in\mathbb{C}S^{2}$ and $\lambda\in\mathbb{C}S^{0}$.
\item For any orthonormal basis~$\left\{ \ket{\psi_{i}}\right\} _{i=1}^{3}$,
we have $1\in\sum_{i=1}^{3}\bar{f}\left(\ket{\psi_{i}}\right)$ and
$\left[1,1\right]-\bar{f}\left(\ket{\psi_{i}}\right)\subseteq\bar{f}\left(\ket{\psi_{j}}\right)+\bar{f}\left(\ket{\psi_{k}}\right)$,
where $i$, $j$, and $k$ are any permutations among $1$, $2$,
and $3$.
\end{itemize}
Conversely, if a function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the above properties, $\bar{f}$ must be an interval-valued
frame function.\end{lemma}\footnote{\yutsung{{Verify the proof...}}}

Since interval-valued probability measures~$\bar{\mu}$ and interval-valued
frame functions~$\bar{f}$ are one-to-one correspondence, studying
1-dimensional projectors mapping to impossible by $\bar{\mu}$ is
the same as studying the points on $\mathbb{C}S^{2}$ mapping to impossible
by $\bar{f}$. These points, denoted by $\bar{f}^{-1}\left(\imposs\right)=\set{\ket{\psi}\in\mathbb{C}^{3}}{\bar{f}(\ket{\psi})=\imposs}$,
play an important role to understand $\bar{f}$. There are 4 different
cases when analyzing $\bar{f}^{-1}\left(\imposs\right)$:
\begin{enumerate}
\item If another pair of states~$\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}\in\bar{f}^{-1}\left(\imposs\right)$\footnote{\yutsung{{Check when }$\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}$ are
complex?}} gives different state of cross product~$\ket{\psi_{0}'}\times\ket{\psi_{1}'}$,
then both $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
might induce $\bar{f}$, but neither $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$
nor $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ in the core of
$\bar{f}$ because $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$ is
inconsistent with $\bar{f}(\proj{\psi_{0}'})=\bar{f}(\proj{\psi_{1}'})=\imposs$
and $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ is inconsistent
with $\bar{f}(\proj{\psi_{0}})=\bar{f}(\proj{\psi_{1}})=\imposs$.
Actually, $\bar{f}$ has an empty core. 
\item If $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
always represent the same state for all $\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}\in\bar{f}^{-1}\left(\imposs\right)$,
then $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is the only pure state
which might induce $\bar{f}$. In another word, if $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, then $\ket{\phi}$ and $\ket{\psi_{0}}\times\ket{\psi_{1}}$
represent the same state. 
\item $\bar{f}^{-1}\left(\imposs\right)$ may contain only one state~$\ket{\psi_{0}}$.
In this case, $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is undefined.
If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$ for some pure
state~$\ket{\phi}$, $\ket{\phi}$ could be any state orthogonal
to $\ket{\psi_{0}}$.
\item $\bar{f}^{-1}\left(\imposs\right)=\emptyset$. If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, we could say nothing about $\ket{\phi}$.
\end{enumerate}
In summary, if we have an idea of $\bar{f}^{-1}\left(\imposs\right)$,
we could have some idea of the pure states in core.

As we did in Fig.~\ref{fig:three-dimensional-4-value}, we could
plot the 1-dimensional projectors in a unit sphere~$S^{2}$~\cite{Hatcher2001,dryden2005}.
Since the state mapping to impossible always contained in a plane,
$\bar{f}^{-1}\left(\imposs\right)$ in the above case 2 - 4 does not
contain any area in $S^{2}$. In contrast, if $\bar{f}^{-1}\left(\imposs\right)$
contains some area in $S^{2}$, it must be the situation of case 1.
Furthermore, the larger the area, the more 1-dimensional projectors
which is inconsistent with any particular Born rule probability measure.
Therefore, the area of $\bar{f}^{-1}\left(\imposs\right)$ could indicate
how much $\bar{f}$ deviate from the Born rule probability measures.

>>>>

Although quantum IVPMs in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-4-value} have empty cores, we might
still think that the probability measure~$\bar{\mu}$ is induced
by some states. In example~\ref{ex:three-dimensional-three-value},
the probability measure~$\bar{\mu}$ might be induced by the state
$\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$.
In example~\ref{ex:three-dimensional-4-value}, the probability measure~$\bar{\mu}$
might be induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$.
In general, each two real 1-dimensional projectors~$\proj{\psi_{0}}$
and $\proj{\psi_{1}}$ mapping to impossible suggests $\bar{\mu}$
might be induced by the pure state~$\ket{\psi_{0}}\times\ket{\psi_{1}}$
in the Hilbert space of dimension 3.

@@

\begin{lemma}A interval-valued frame function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the following properties:
\begin{itemize}
\item $\bar{f}\left(\lambda\ket{\psi}\right)=\bar{f}\left(\ket{\psi}\right)$
for any $\ket{\psi}\in\mathbb{C}S^{2}$ and $\lambda\in\mathbb{C}S^{0}$.
\item For any orthonormal basis~$\left\{ \ket{\psi_{i}}\right\} _{i=1}^{3}$,
we have $1\in\sum_{i=1}^{3}\bar{f}\left(\ket{\psi_{i}}\right)$ and
$\left[1,1\right]-\bar{f}\left(\ket{\psi_{i}}\right)\subseteq\bar{f}\left(\ket{\psi_{j}}\right)+\bar{f}\left(\ket{\psi_{k}}\right)$,
where $i$, $j$, and $k$ are any permutations among $1$, $2$,
and $3$.
\end{itemize}
Conversely, if a function~$\bar{f}:\mathbb{C}S^{2}\rightarrow\mathscr{I}$
satisfies the above properties, $\bar{f}$ must be an interval-valued
frame function.\end{lemma}\footnote{\yutsung{Verify the proof...}}

\begin{definition}For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow: 
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi}\times\ket{\phi}$ need not be normalized
as usual. We will use $\ket{\psi}\times\ket{\phi}$ and $\frac{\ket{\psi}\times\ket{\phi}}{\left\Vert \ket{\psi}\times\ket{\phi}\right\Vert }$
interchanged when there is no confusion.\qed\end{definition}

\begin{lemma}Given $\ket{\psi}$, $\ket{\phi}$ and $\ket{\varphi}\in\mathbb{C}^{3}$
such that $\ket{\psi}$ is not parallel to $\ket{\phi}$. Then, $\ket{\varphi}$
is parallel to $\ket{\psi}\times\ket{\phi}$ if and only if $\ip{\psi}{\varphi}=\ip{\phi}{\varphi}=0$.\end{lemma}
\begin{proof}
To prove that $\ket{\varphi}$ is parallel to $\ket{\psi}\times\ket{\phi}\Rightarrow\ip{\psi}{\varphi}=\ip{\phi}{\varphi}=0$,
we can just verify 
\begin{eqnarray*}
\ip{\psi}{\psi\times\phi} & = & \ip{\psi\times\phi}{\psi}^{*}=\left[\left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}\right]^{*}=0\textrm{ ,}\\
\ip{\phi}{\psi\times\phi} & = & \ip{\psi\times\phi}{\phi}^{*}=\left[\left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}\right]^{*}=0\textrm{ .}
\end{eqnarray*}

To prove the other direction, consider the linear transformation~$T:\mathbb{C}^{3}\rightarrow\mathbb{C}^{2}$
defined by $T\left(\ket{\varphi}\right)=\begin{pmatrix}\ip{\psi}{\varphi}\\
\ip{\phi}{\varphi}
\end{pmatrix}$. Its rank equation~\cite{FraleighBeauregard1995} implies
\begin{eqnarray*}
\dim\left(\ker T\right) & = & \mathrm{nullity}\left(T\right)\\
 & = & \dim\left(\mathrm{domain}T\right)-\mathrm{rank}\left(T\right)\\
 & = & 3-2=1
\end{eqnarray*}
By the other direction of the proof, we have already known $\ket{\psi}\times\ket{\phi}\in\ker T$.
Therefore, $\ket{\varphi}$ is parallel to $\ket{\psi}\times\ket{\phi}$.
\end{proof}
Since interval-valued probability measures~$\bar{\mu}$ and interval-valued
frame functions~$\bar{f}$ are one-to-one correspondence, studying
1-dimensional projectors mapping to impossible by $\bar{\mu}$ is
the same as studying the points on $\mathbb{C}S^{2}$ mapping to impossible
by $\bar{f}$. These points, denoted by $\bar{f}^{-1}\left(\imposs\right)=\set{\ket{\psi}\in\mathbb{C}^{3}}{\bar{f}(\ket{\psi})=\imposs}$,
play an important role to understand $\bar{f}$. There are 4 different
cases when analyzing $\bar{f}^{-1}\left(\imposs\right)$:
\begin{enumerate}
\item If another pair of states~$\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}\in\bar{f}^{-1}\left(\imposs\right)$
gives different state of cross product~$\ket{\psi_{0}'}\times\ket{\psi_{1}'}$,
then both $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
might induce $\bar{f}$, but neither $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$
nor $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ in the core of
$\bar{f}$ because $f_{\ket{\psi_{0}}\times\ket{\psi_{1}}}^{B}$ is
inconsistent with $\bar{f}(\proj{\psi_{0}'})=\bar{f}(\proj{\psi_{1}'})=\imposs$
and $f_{\ket{\psi_{0}'}\times\ket{\psi_{1}'}}^{B}$ is inconsistent
with $\bar{f}(\proj{\psi_{0}})=\bar{f}(\proj{\psi_{1}})=\imposs$.
Actually, $\bar{f}$ has an empty core. 
\item If $\ket{\psi_{0}}\times\ket{\psi_{1}}$ and $\ket{\psi_{0}'}\times\ket{\psi_{1}'}$
always represent the same state for all $\ket{\psi_{0}'}$ and $\ket{\psi_{1}'}\in\bar{f}^{-1}\left(\imposs\right)$,
then $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is the only pure state
which might induce $\bar{f}$. In another word, if $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, then $\ket{\phi}$ and $\ket{\psi_{0}}\times\ket{\psi_{1}}$
represent the same state. 
\item $\bar{f}^{-1}\left(\imposs\right)$ may contain only one state~$\ket{\psi_{0}}$.
In this case, $\ket{\psi_{0}}\times\ket{\psi_{1}}$ is undefined.
If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$ for some pure
state~$\ket{\phi}$, $\ket{\phi}$ could be any state orthogonal
to $\ket{\psi_{0}}$.
\item $\bar{f}^{-1}\left(\imposs\right)=\emptyset$. If $f_{\phi}^{B}\in\mathrm{core}\left(\bar{f}\right)$
for some pure state~$\ket{\phi}$, we could say nothing about $\ket{\phi}$.
\end{enumerate}
In summary, if we have an idea of $\bar{f}^{-1}\left(\imposs\right)$,
we could have some idea of the pure states in core.

As we did in Fig.~\ref{fig:three-dimensional-4-value}, we could
plot the 1-dimensional projectors in a unit sphere~$S^{2}$~\cite{Hatcher2001,dryden2005}.
Since the state mapping to impossible always contained in a plane,
$\bar{f}^{-1}\left(\imposs\right)$ in the above case 2 - 4 does not
contain any area in $S^{2}$. In contrast, if $\bar{f}^{-1}\left(\imposs\right)$
contains some area in $S^{2}$, it must be the situation of case 1.
Furthermore, the larger the area, the more 1-dimensional projectors
which is inconsistent with any particular Born rule probability measure.
Therefore, the area of $\bar{f}^{-1}\left(\imposs\right)$ could indicate
how much $\bar{f}$ deviate from the Born rule probability measures.\footnote{Notice that we only discuss only discuss real 1-dimensional projectors
in the Hilbert space of dimension 3. In general, we could discuss
all 1-dimensional projectors in any dimension. However, restricting
to $\mathbb{R}^{3}$ enable us to use cross product which will simplify
our discussion a lot, and $\mathbb{R}^{3}$ is easy to visualize and
imagine. Also, if a quantum IVPM~$\bar{\mu}$ has an empty core when
restricting to $\mathbb{R}^{3}$, $\bar{\mu}$ has an empty core on
the original space as well. So restricting on real 1-dimensional projectors
still provides valid results.}
